---
title: "Carter Athens"
author: "Carter Athens"
date: "April 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library("tidyverse")
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)
names(per)
```

Part one: Accident data


Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)? 
The most accidents happen on Saturday. Sunday and Friday also have a lot more accidents than the rest of the week.
```{r}
ggplot(acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", binwidth = .5) + scale_x_continuous(breaks=seq(1,7,1)) + xlab("Day of Week (1 is Sunday, 7 is Saturday)") + ylab("Accidents") + ggtitle("Accidents Per Day of Week")
```


What time of the day do accidents happen (use variable HOUR)?
Most accidents happen around hour 18. This is because most people are getting off work.
```{r}
acc <- acc[!(acc$HOUR == "99"),] # The HOUR column has some wrong values so I removed them
ggplot(acc, aes(x=HOUR)) + geom_histogram(color="black", binwidth = .5) + scale_x_continuous(breaks=seq(0,23,1)) + xlab("Hour") + ylab("Accidents") + ggtitle("Accidents Per Hour")
```


What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
8,769 accidents involved at least one drunk driver.
```{r}
numAccWithDrunkDriver <- filter(acc, DRUNK_DR > 0)
dim(numAccWithDrunkDriver)
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```