---
title: "team"
author: "Carter Athens and Max Ruehle"
date: "April 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)
names(per)
```

Part one: Accident data


Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)? 
The most accidents happen on Saturday. Sunday and Friday also have a lot more accidents than the rest of the week.
```{r}
ggplot(acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", binwidth = .5) + scale_x_continuous(breaks=seq(1,7,1)) + xlab("Day of Week (1 is Sunday, 7 is Saturday)") + ylab("Accidents") + ggtitle("Accidents Per Day of Week")
```


What time of the day do accidents happen (use variable HOUR)?
Most accidents happen around hour 18. This is because most people are getting off work.
```{r}
acc <- acc[!(acc$HOUR == "99"),] # The HOUR column has some wrong values so I removed them
ggplot(acc, aes(x=HOUR)) + geom_histogram(color="black", binwidth = .5) + scale_x_continuous(breaks=seq(0,23,1)) + xlab("Hour") + ylab("Accidents") + ggtitle("Accidents Per Hour")
```


What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
8,769 accidents involved at least one drunk driver.
```{r}
numAccWithDrunkDriver <- filter(acc, DRUNK_DR > 0)
dim(numAccWithDrunkDriver)
```


Part two: Connecting data


Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them. Join accident and driver table (work out which variable(s) to use). Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
drivers <- per %>% filter(PER_TYP == 1)

drivers %>% select(-c(HOUR)) %>% left_join(acc, by="ST_CASE") %>% group_by(DAY_WEEK) %>% summarise(count = n()) %>% ggplot(aes(x=DAY_WEEK)) + scale_x_continuous(breaks=seq(1,7,1)) + geom_bar(aes(weight=count))
# There are more accidents on Sunday, Friday, and Saturday.

drivers %>% select(-c(HOUR)) %>% left_join(acc, by="ST_CASE") %>% group_by(HOUR) %>% summarise(count = n()) %>% ggplot(aes(x=HOUR)) + scale_x_continuous(breaks=seq(0,23,1)) + geom_bar(aes(weight=count))
# Most accidents happen during rush hour when people are getting off work.

drivers %>% select(-c(HOUR)) %>% left_join(acc, by="ST_CASE") %>% group_by(SEX) %>% summarise(count = n()) %>% ggplot(aes(x=SEX)) + geom_bar(aes(weight=count))
# Men are involved in accidents significantly more often than women.
```